---
output: github_document
---

## Create a map of Australia and Map Data Collection Points

This is a simple RMD file to illustrate how to use
[_rnaturalearth_](https://github.com/ropenscilabs/rnaturalearth), [_simple features_](https://cran.r-project.org/web/packages/sf/vignettes/sf1.html) and
_ggplot2_ to create a map of Australia and plot data collection points on it.

### Setup

To do this you'll need a few packages from CRAN:

```{r pkg-load, eval=TRUE, messages=FALSE, warning=FALSE}
#devtools::install_github("ropenscilabs/rnaturalearthdata")
#devtools::install_github("tidyverse/ggplot2")

library(tidyverse)
library(rnaturalearth)
library(raster)
library(sf)

```

## Get the Data to Create Our Map

### Add a Shapefile of Australia

This is our base layer, Australia, of the map from (Naturalearth.com)[https://naturalearth.com/]

```{r, australia}

oz_shape <- rnaturalearth::ne_states(geounit = "australia",
                                     returnclass = "sf")

plot(oz_shape)

```

However, it includes several islands and ocean that are not of interest to us.
To fix this, crop it down to just the mainland plus Tasmania and remove Jervis Bay so that the labels on the final map product are cleaner.

```{r crop_shape}

oz_shape <- st_intersection(oz_shape, 
                st_set_crs(
                  st_as_sf(
                    as(
                      raster::extent(114,
                                     155,
                                     -45,
                                     -9),
                      "SpatialPolygons")),
                  st_crs(oz_shape)))

oz_shape <- oz_shape[oz_shape$woe_name != "Jervis Bay", ]

plot(oz_shape)
```

## Import Collection Point Data

Assuming the "Geocode_319 samples.csv" file is in the same working directory as this script.

```{r collection_data}
point_data <- read_csv("Geocode_319 samples.csv")
point_data$State <- as.factor(point_data$State)
```

## Plot using `ggplot2`

Plot the final combination of data with the Naturalearth data Australia and state outlines with the sampling location points overplayed and colour-coded by the state where they were sampled.

```{r plot, eval=TRUE}

oz <- ggplot(oz_shape) +
  geom_sf(fill = "white") +
  geom_text(
  data = oz_shape,
  aes(x = longitude,
  y = latitude,
  label = abbrev),
  size = 2.5,
  hjust = 1
  ) +
  geom_point(data = point_data,
  aes(x = Longitude,
  y = Latitude,
  colour = State),
  alpha = 0.6,
  size = 2) +
  theme_bw() +
  xlab("Longitude") +
  ylab("Latitude") +
  theme(legend.position = "none") +
  coord_sf()

oz
```

## Save the graph

Export at 500 dpi for publication with a width 190mm for a 2-column width in an Elsevier publication.

```{r, eval = FALSE}
ggsave("Australia_Map.tiff", width = 190, units = "mm", dpi = 500)
```

# Appendix

```{r sessionInfo, eval=TRUE, echo=FALSE}
sessionInfo()
```
