---
output: github_document
---
```{r opts}
knitr::opts_chunk$set(
  fig.path = "figures/README_files"
)
```

## Creating a Map of Australia, Adding Data for Sample Collection Points

This is a simple RMD file to illustrate how to use
[_rnaturalearth_](https://github.com/ropenscilabs/rnaturalearth), [_simple features_](https://cran.r-project.org/web/packages/sf/vignettes/sf1.html) and
_ggplot2_ to create a map of Australia and plot data collection points on it.

### Setup

To do this you'll need a few packages from CRAN:

```{r pkg-load, eval=TRUE, messages=FALSE, warning=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}

library("pacman")
p_load(tidyverse, rnaturalearth, rnaturalearthdata, raster, sf, rgeos, lwgeom)
if (!require("rnaturalearthhires")) {
  p_install_gh("ropenscilabs/rnaturalearthhires")
}
```

### Add a Shapefile of Australia

This is our base layer, Australia, of the map from (Naturalearth.com)[https://naturalearth.com/].

```{r australia}
oz_sf <- rnaturalearth::ne_states(geounit = "australia",
                                  returnclass = "sf")

plot(oz_sf)
```

However, it includes several islands and ocean that are not of interest to us.
To fix this, crop it down to just the mainland plus Tasmania and remove Jervis Bay so that the labels on the final map product are cleaner.

```{r crop_sf}
oz_sf <- st_intersection(oz_sf,
                         st_set_crs(st_as_sf(as(
                           raster::extent(114,
                                          155,
                                          -45,
                                          -9),
                           "SpatialPolygons"
                         )),
                         st_crs(oz_sf)))

oz_sf <- oz_sf[oz_sf$abbrev != "J.B.T.",]

plot(oz_sf)
```

## Import Collection Point Data

If you have your own data in a .csv file, import it using `read_csv()`.
These are just random points that I generated and saved as a .csv file for this work to illustrate how to add points to a map.

```{r collection_data, eval=TRUE}
point_data <- read_csv("sample_points.csv",
                       col_types = cols(
                         col_double(),
                         col_double(),
                         col_factor()
                       ))
```

## Plot Using ggplot2

Plot the final combination of data with the Naturalearthdata Australia data and state outlines with the sampling location points overplayed and colour-coded by the state where they were sampled.

```{r plot, eval=TRUE}
oz <- ggplot(oz_sf) +
  geom_sf(fill = "white") +
  geom_text(
    data = oz_sf,
    aes(x = longitude,
        y = latitude,
        label = abbrev),
    size = 2.5,
    hjust = 1
  ) +
  geom_point(data = point_data,
             aes(x = Longitude,
                 y = Latitude,
                 colour = class),
             size = 2) +
  theme_bw() +
  scale_colour_brewer(type = "Qualitative", palette = "Set1") +
  xlab("Longitude") +
  ylab("Latitude") +
  coord_sf()

oz
```

## Save the Graph

Export at 500 DPI for publication with a width 190mm for a 2-column width figure.

```{r, eval=FALSE}
ggsave("Australia_Map.tiff",
       width = 190,
       units = "mm",
       dpi = 500)
```

## Meta

### Code of Conduct

Please note that the R_map_with_points_Oz project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

### R Session Information

```{r session_info, eval=TRUE, echo=FALSE}
sessionInfo()
```
